---
title: "Health Expenditure"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width=10, fig.height=8)

library(flexdashboard)
library(ggplot2)
library(dplyr)
library(terra)
library(stars)
library(sf)
# install.packages("terra")
library(terra)
library("raster")
library(raster)
library(rgeos)

library(tmap)
library(ggplot2)
library(wesanderson)
```

Column {.tabset}
-----------------------------------------


### Boxplot Analysis of Mortality 

```{r boxplot_death, echo = FALSE}
#health_exp_cases <- readRDS("df_health_expenditure_cases.RDS")

health_exp_deaths <- readRDS("df_health_expenditure_deaths.RDS")

health_exp_deaths$`Deaths per 100000` <- health_exp_deaths$deaths_per_capita * 100000

health_exp_deaths <- health_exp_deaths %>%
  filter(!is.na(health_exp_deaths$df_health_conf_categories))

box_1 <- ggplot(health_exp_deaths, aes(x = df_health_conf_categories, y = `Deaths per 100000`, fill = df_health_conf_categories)) +  geom_bar(stat = "identity", na.rm=T) + scale_fill_manual(values = wes_palette("Zissou1")) + labs(y="Deaths per 100000",
       x="Healthcare Expenditure by % of GDP",   
       title="COVID mortality rate and Healthcare expenditure by % of GDP") +
  theme(legend.position="none")

box_1

```
**_NOTE:_** The ANOVA test found there was a significant difference between these categories in impacting COVID mortality. The post-hoc Tukey test found significant differences for High-Low, Very high-Low and High-Moderate categories. 


```{r anova_mortality_health, include=FALSE}
##ANOVA/KRUSKAl ##2 -- 3 categories are significant 
one.way_mort <- aov(health_exp_deaths$deaths_per_capita ~ health_exp_deaths$df_health_conf_categories, data = health_exp_deaths)
summary(one.way_mort)

tukey.test_healthexp <- TukeyHSD(one.way_mort)
tukey.test_healthexp
```


### Visual Scatterplot of Mortality
```{r scatterplot_death, echo = FALSE}
scatter_1 <- ggplot(health_exp_deaths, aes(x = `2021`, y = `Deaths per 100000`)) + geom_point(aes(col=df_health_conf_categories)) +
  labs(y="Deaths per 100000",
       x="Healthcare Expenditure by % of GDP",   
       title="COVID mortality rate and Healthcare expenditure by % of GDP") +
  guides(col=guide_legend("Healthcare Expenditure"))

scatter_1

```


### Interactive World map of COVID mortality
```{r tmap_death, echo = FALSE}
# Joining the tmap world data and our mortality data using common iso3 name coloumn
data(World)
death_geom <- left_join(health_exp_deaths,World, by = c("iso3c" = "iso_a3"))

# Converting the new dataset into a shape file  
death_geom <- st_as_sf(death_geom)

# Removing NA's from the sf-df: the coloumn gd was missing data when there were no coordinates.
death_geom_clean <- death_geom %>%
  filter(!is.na(gdp_cap_est))
# Changing coloumn name for aestheitc reasons
death_geom_clean$`Healthcare Expenditure` <- death_geom_clean$df_health_conf_categories

# Setting up a gradient color for mortality

tmap_1 <- tm_shape(death_geom_clean) + 
    tm_polygons("Healthcare Expenditure", palette = "Blues", title = "Healthcare expenditure", contrast =0.5, clustering = FALSE) + tm_text("iso3c", size = 0.5) + 
    tm_shape(death_geom_clean) +
    tm_bubbles("Deaths per 100000",
               border.col = "black", border.alpha = .5, style="fixed",
               breaks=c(0, 50,100,150,200,Inf),
               col="Deaths per 100000",
               n = 6,
               clustering = FALSE,
 title.size="Mortality per 100000", title.col="COVID Mortality") +
    tm_facets(as.layers = TRUE)

# view map with default view options
tmap_mode("view")
tmap_1

```
> **_NOTE:_**  World Map shows countries with varying COVID mortality along with healthcare expenditure data.  

Column {.tabset}
-----------------------------------------


### Boxplot Analysis of Confirmed Cases 

```{r boxplot_cases, echo = FALSE}

health_exp_cases <- readRDS("df_health_expenditure_cases.RDS")

health_exp_cases$`Cases per 100000` <- health_exp_cases$cases_per_capita * 100000

health_exp_cases <- health_exp_cases %>%
  filter(!is.na(health_exp_cases$df_health_conf_categories))

box_2 <- ggplot(health_exp_cases, aes(x=factor(df_health_conf_categories), y=`Cases per 100000`,  fill=factor(df_health_conf_categories)))  +  geom_bar(stat = "identity", na.rm=T) + scale_fill_manual(values = wes_palette("Zissou1")) + labs(y="Cases per 100000",
       x="Healthcare Expenditure by % of GDP",   
       title="COVID confirmed cases rate and Healthcare expenditure by % of GDP") +
  theme(legend.position="none")

box_2
```
**_NOTE:_** The ANOVA test found there was a significant difference between these categories in impacting COVID case count incidence. The post-hoc Tukey test found significant differences for High-Low, Very high-Low and High-Moderate categories and Very high-Moderate categories. 

```{r anova_cases_health, include=FALSE}
##ANOVA/KRUSKAl ##2 -- 3 categories are significant 
one.way_cases <- aov(health_exp_cases$cases_per_capita ~ health_exp_cases$df_health_conf_categories, data = health_exp_cases)
summary(one.way_cases)

tukey.test_healthexp_cases <- TukeyHSD(one.way_cases)
tukey.test_healthexp_cases
```

### Visual Scatterplot of Confirmed Cases
```{r scatterplot_cases, echo = FALSE}
scatter_2 <- ggplot(health_exp_cases, aes(x = `2021`, y = `Cases per 100000`)) + geom_point(aes(col=df_health_conf_categories)) +
  labs(y="Cases per 100000",
       x="Healthcare Expenditure by % of GDP",   
       title="COVID confirmed cases rate and Healthcare expenditure by % of GDP") +
  guides(col=guide_legend("Healthcare Expenditure"))

scatter_2
```
  

### Interactive World Map of COVID cases
```{r tmap_cases, echo = FALSE}
# Joining the tmap world data and our mortality data using common iso3 name coloumn
data("World")
cases_geom <- left_join(health_exp_cases,World, by = c("iso3c" = "iso_a3"))

# Converting the new dataset into a shape file  
cases_geom <- st_as_sf(cases_geom)

# Removing NA's from the sf-df: the coloumn gd was missing data when there were no coordinates.
cases_geom_clean <- cases_geom %>%
  filter(!is.na(gdp_cap_est))
# Changing coloumn name for aestheitc reasons
cases_geom_clean$`Healthcare Expenditure` <- cases_geom_clean$df_health_conf_categories

# Setting up a gradient color for mortality

tmap_2 <- tm_shape(cases_geom_clean) + 
    tm_polygons("Healthcare Expenditure", palette = "Blues", title = "Healthcare expenditure", contrast =0.5, clustering = FALSE) + tm_text("iso3c", size = 0.5) + 
    tm_shape(cases_geom_clean) +
    tm_bubbles("Cases per 100000",
               border.col = "black", border.alpha = .5, style="fixed",
               breaks=c(0, 450,4000,10000,30000),
               col="Cases per 100000",
               n = 6,
               clustering = FALSE,
 title.size="Cases per 100000", title.col="COVID Cases") +
    tm_facets(as.layers = TRUE)

# view map with default view options
tmap_mode("view")
tmap_2

```
> **_NOTE:_**  World Map shows countries with varying COVID case count along with healthcare expenditure data.  


